Temporal Dynamics of Normalization Reweighting

### 2023-03-19 manuscript.Rmd
Minor edits for submission to JoV

---
title: 'Temporal Dynamics of Normalization Reweighting'
author: | 
  | Daniel H. Baker, Daniela Marinova, Richard Aveyard, Lydia J. Hargreaves, Alice Renton,  
  | Ruby Castellani, Phoebe Hall, Miriam Harmens, Georgia Holroyd, Beth Nicholson, 
  | Emily L. Williams, Hannah M. Hobson & Alex R. Wade
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

# Significance statement

We investigated the timecourse of a novel form of neural plasticity, normalization reweighting, using steady-state EEG in humans. We find increases in suppression during the first 2-5 seconds of stimulation, with similar dynamics for monocular and dichoptic presentation. The effects are present in primary visual cortex, and were not mediated by autistic traits. These findings demonstrate that reweighting builds up rapidly after stimulus onset, consistent with a process that adjusts the sensory response to the local environment. Normalization reweighting also has potential as a biomarker of neural function both in humans and in animal models of disease.

@@

Experiment 1 was completed by 100 adult participants (32 male, 68 female; mean age 21.9) in early 2015, and first reported by @Vilidaite2018. Here we reanalysed the dataset, and report the results of masking conditions not previously published. Experiment 2 was completed by 100 adult participants (23 male, 74 female, 3 other/not stated; mean age 22.1) in early 2022. Experiment 3 was completed by 10 adults (2 male, 8 female) with a clinical diagnosis of autism, and 10 control participants who were closely matched for age (means of `r t_age$estimate[1]` and `r t_age$estimate[2]`, *t* = `r format(round(abs(t_age$statistic),digits=2))`, *df* = `r t_age$parameter`, *p* = `r format(round(abs(t_age$p.value),digits=2))`) and exactly matched for gender. Procedures in Experiments 1 and 2 were approved by the ethics committee of the Department of Psychology at the University of York. Procedures for Experiment 3 were approved by the ethics committee of the York Neuroimaging Centre. All participants provided written informed consent.

>>>>

---
title: 'Temporal Dynamics of Normalization Reweighting'
author: | 
  | Daniel H. Baker, Daniela Marinova, Richard Aveyard, Lydia J. Hargreaves, Alice Renton,  
  | Ruby Castellani, Phoebe Hall, Miriam Harmens, Georgia Holroyd, Beth Nicholson, 
  | Emily L. Williams, Hannah M. Hobson & Alex R. Wade
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
    extra_dependencies: ["lineno"]
  bookdown::word_document2:
    fig_caption: yes
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

\linenumbers

@@

@@

Experiment 1 was completed by 100 adult participants (32 male, 68 female; mean age 21.9) in early 2015, and first reported by @Vilidaite2018. Here we reanalysed the dataset, and report the results of masking conditions not previously published. Experiment 2 was completed by 100 adult participants (23 male, 74 female, 3 other/not stated; mean age 22.1) in early 2022. Experiment 3 was completed by 10 adults (2 male, 8 female) with a clinical diagnosis of autism, and 10 control participants who were closely matched for age (means of `r t_age$estimate[1]` and `r t_age$estimate[2]`, *t* = `r format(round(abs(t_age$statistic),digits=2))`, *df* = `r t_age$parameter`, *p* = `r format(round(abs(t_age$p.value),digits=2))`) and exactly matched for gender. Procedures in Experiments 1 and 2 were approved by the ethics committee of the Department of Psychology at the University of York. Procedures for Experiment 3 were approved by the ethics committee of the York Neuroimaging Centre. All participants provided written informed consent, and procedures were consistent with the Declaration of Helsinki.

### 2023-03-19 manuscript.tex
Minor edits for submission to JoV

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{lineno}
\linenumbers
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Temporal Dynamics of Normalization Reweighting},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

@@

\hypertarget{significance-statement}{%
\section{Significance statement}\label{significance-statement}}

We investigated the timecourse of a novel form of neural plasticity, normalization reweighting, using steady-state EEG in humans. We find increases in suppression during the first 2-5 seconds of stimulation, with similar dynamics for monocular and dichoptic presentation. The effects are present in primary visual cortex, and were not mediated by autistic traits. These findings demonstrate that reweighting builds up rapidly after stimulus onset, consistent with a process that adjusts the sensory response to the local environment. Normalization reweighting also has potential as a biomarker of neural function both in humans and in animal models of disease.

@@

Experiment 1 was completed by 100 adult participants (32 male, 68 female; mean age 21.9) in early 2015, and first reported by Vilidaite et al. (2018). Here we reanalysed the dataset, and report the results of masking conditions not previously published. Experiment 2 was completed by 100 adult participants (23 male, 74 female, 3 other/not stated; mean age 22.1) in early 2022. Experiment 3 was completed by 10 adults (2 male, 8 female) with a clinical diagnosis of autism, and 10 control participants who were closely matched for age (means of 21.8 and 22, \emph{t} = 0.18, \emph{df} = 18, \emph{p} = 0.86) and exactly matched for gender. Procedures in Experiments 1 and 2 were approved by the ethics committee of the Department of Psychology at the University of York. Procedures for Experiment 3 were approved by the ethics committee of the York Neuroimaging Centre. All participants provided written informed consent.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Temporal Dynamics of Normalization Reweighting},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\linenumbers

@@

@@

Experiment 1 was completed by 100 adult participants (32 male, 68 female; mean age 21.9) in early 2015, and first reported by Vilidaite et al. (2018). Here we reanalysed the dataset, and report the results of masking conditions not previously published. Experiment 2 was completed by 100 adult participants (23 male, 74 female, 3 other/not stated; mean age 22.1) in early 2022. Experiment 3 was completed by 10 adults (2 male, 8 female) with a clinical diagnosis of autism, and 10 control participants who were closely matched for age (means of 21.8 and 22, \emph{t} = 0.18, \emph{df} = 18, \emph{p} = 0.86) and exactly matched for gender. Procedures in Experiments 1 and 2 were approved by the ethics committee of the Department of Psychology at the University of York. Procedures for Experiment 3 were approved by the ethics committee of the York Neuroimaging Centre. All participants provided written informed consent, and procedures were consistent with the Declaration of Helsinki.
